---
title: "ANALYSIS"
---

## Overview of the Data

Our analysis will be based on the data set Economic Confidence Index set published by TÜİK. The ECI reflects the general sentiment towards the economy and is composed of various sectoral indices. The data set spans from January 2011 to October 2024 and provides information on the following indices:

-   **Economic Confidence Index (ECI)**: Represents the overall economic outlook by combining data from multiple sectors.

-   **Consumer Confidence Index (CCI)**: This shows how consumers perceive the current and future state of the economy, their financial situations, and their spending tendencies.

-   **Real Sector Confidence Index (RSCI)**: Focuses on the manufacturing industry, providing insights into production levels, new orders, and business expectations.

-   **Service Confidence Index (SCI)**: Highlights the performance and expectations within the service sector.

-   **Retail Trade Confidence Index (RTCI)**: Focuses on sales trends, stock levels, and business activity in retail.

-   **Construction Confidence Index (CNCI)**: Tracks sentiment in the construction industry, including current order books and employment plans.

This data set provides a comprehensive view of the Turkish economy through sectoral lenses, making it useful for analyzing trends, comparing sector performance, and evaluating economic resilience over time.

In the data set rows represent the data of that year with the belongings month, and the columns represent the different indices explained above. The column names are: "Year", "Month", "Economic_Confidence\_ Index", "Consumer_Confidence_Index", "Real_Sector_Confidence_Index", "Services_Confidence_Index", "Retail_Trade_Confidence_Index", "Construction_Confidence_Index."

In the below, you can see the distributions and value types of the data. We found the variable type by using the str() function, and the distribution by looking at the histograms of it. You can see the all details about the data in the following table:

+--------------------------------------+-----------------------------------------------------------------------------------------------------------------------------+----------------------+-------------------------------------------------------------------------------------------+
| Variable                             | Meaning                                                                                                                     | Type                 | Value Distribution                                                                        |
+======================================+=============================================================================================================================+======================+===========================================================================================+
| Year, Month                          | The date when the data was recorded.                                                                                        | Numeric              | This column will contain a range of dates from 2011 to 2024, likely at monthly intervals. |
|                                      |                                                                                                                             |                      |                                                                                           |
|                                      |                                                                                                                             | (Discrete)           |                                                                                           |
+--------------------------------------+-----------------------------------------------------------------------------------------------------------------------------+----------------------+-------------------------------------------------------------------------------------------+
| Economic Confidence Index (ECI)      | The composite index that represents overall economic confidence.                                                            | Numeric (Continuous) | Normal Distribution                                                                       |
+--------------------------------------+-----------------------------------------------------------------------------------------------------------------------------+----------------------+-------------------------------------------------------------------------------------------+
| Consumer Confidence Index (CCI)      | A measure of consumers' confidence about their financial situations and expectations for the economy.                       | Numeric (Continuous) | Normal Distribution                                                                       |
+--------------------------------------+-----------------------------------------------------------------------------------------------------------------------------+----------------------+-------------------------------------------------------------------------------------------+
| Real Sector Confidence Index (RSCI)  | Reflects confidence within the real sector, such as production levels, investments, and employment trends in manufacturing. | Numeric (Continuous) | Normal Distribution                                                                       |
+--------------------------------------+-----------------------------------------------------------------------------------------------------------------------------+----------------------+-------------------------------------------------------------------------------------------+
| Service Confidence Index (SCI)       | Reflects the confidence in the service sector, considering factors like business activity and expectations.                 | Numeric (Continuous) | Normal Distribution                                                                       |
+--------------------------------------+-----------------------------------------------------------------------------------------------------------------------------+----------------------+-------------------------------------------------------------------------------------------+
| Retail Trade Confidence Index (RTCI) | Represents the sentiment in retail trade, including expectations of sales and stock levels.                                 | Numeric (Continuous) | Normal Distribution                                                                       |
+--------------------------------------+-----------------------------------------------------------------------------------------------------------------------------+----------------------+-------------------------------------------------------------------------------------------+
| Construction Confidence Index (CNCI) | Reflects confidence in the construction sector, focusing on current orders and future business activity.                    | Numeric (Continuous) | Normal Distribution                                                                       |
+--------------------------------------+-----------------------------------------------------------------------------------------------------------------------------+----------------------+-------------------------------------------------------------------------------------------+

## Time Series Analysis

### Seasonal Analysis

```{r message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Show the code"

# Load necessary libraries
library(readr)
library(dplyr)
library(ggplot2)
library(knitr)
library(tidyr)

# Import the dataset
data <- read.csv("Cleaned_Economic_Confidence_Index_Dataset.csv")

# Exclude 'Year' and 'Month' columns (~ AI ~)
columns_to_analyze <- setdiff(names(data), c("Year", "Month"))

# Perform seasonal analysis for all columns except 'Year' and 'Month'
seasonal_analysis <- data %>%
  group_by(Month) %>%
  summarise(across(all_of(columns_to_analyze), ~ mean(.x, na.rm = TRUE)))

# Reshape the data for faceted plotting, We learned the use of pivot_longer() function from DataCamp courses.
melted_data <- seasonal_analysis %>%
  pivot_longer(cols = -Month, names_to = "Index", values_to = "Seasonal_Average")

# Set the order of facets to match the order in the Excel file (~ AI ~)
melted_data$Index <- factor(melted_data$Index, levels = columns_to_analyze)

# Plot using faceting to compare seasonal averages of different indices
ggplot(melted_data, aes(x = Month, y = Seasonal_Average)) +
  geom_line(aes(group = Index), color = "blue") +
  geom_point(color = "red") +
  facet_wrap(~ Index, scales = "free_y") +
  labs(title = "Seasonal Analysis of All Indices",
       x = "Month",
       y = "Seasonal Average") +
  scale_x_continuous(breaks = seq(2, 12, by = 2)) +
  theme_minimal()

```

*Initially, we plotted each variable separately. However, we realized that this approach was inefficient, both in terms of space and clarity, as it required extensive code for minimal insight. To address this, we adopted faceting, which allowed us to present all variables in a single, well-organized visualization. This approach significantly improved the efficiency of our analysis, saving both time and space while enhancing interpretability.*

*Our methodology was primarily guided by lecture materials and DataCamp assignments, with occasional assistance from AI tools. We annotated sections of code where AI provided input with the comment "\~ AI \~" for transparency.*

### Yearly Comparisons

```{r message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Show the code"

# Load necessary libraries
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)

# Import the dataset
data <- read.csv("Cleaned_Economic_Confidence_Index_Dataset.csv")

# Exclude 'Year' and 'Month' columns (~ AI ~)
columns_to_analyze <- setdiff(names(data), c("Year", "Month"))

# Perform yearly analysis for all columns except 'Year' and 'Month'
yearly_analysis <- data %>%
  group_by(Year) %>%
  summarise(across(all_of(columns_to_analyze), ~ mean(.x, na.rm = TRUE)))

# Reshape the data for faceted plotting
melted_data <- yearly_analysis %>%
  pivot_longer(cols = -Year, names_to = "Index", values_to = "Yearly_Average")

# Set the order of facets to match the order in the Excel file (~ AI ~)
melted_data$Index <- factor(melted_data$Index, levels = columns_to_analyze)

# Plot using faceting with vertical x-axis labels for yearly analysis
ggplot(melted_data, aes(x = Year, y = Yearly_Average)) +
  geom_line(aes(group = Index), color = "blue") +
  geom_point(color = "red") +
  facet_wrap(~ Index, scales = "free_y") +
  labs(title = "Yearly Analysis of All Indices",
       x = "Year",
       y = "Yearly Average") +
  scale_x_continuous(breaks = unique(data$Year)) + # Ensure all years are shown (~ AI ~)
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1) # Rotate x-axis labels
  )



```
